\begin{center} \bf \sffamily \LARGE Prisoner's dilemma with stochastic payoffs\\[-1cm] \end{center}

### 2021-03-16 main.tex
a commit with some writting

\documentclass[11pt]{article}
\usepackage{times}
\usepackage{amsmath,amsthm,amssymb,setspace,enumitem,epsfig,titlesec,verbatim,color,array,eurosym,multirow}
\usepackage[sort&compress]{natbib}
\usepackage[footnotesize,bf]{caption}
\usepackage[top=2.2cm,left=2.4cm,right=2.4cm,bottom=3cm]{geometry}
\smallskip % Erlaubt kleine Abstaende zwischen Paragraphen, falls es dem Seitenlayout hilft
\renewcommand{\baselinestretch}{1.3}

@@

%\section*{Extended Data Figures}

\section{Games dynamics between reactive strategies}

Consider two players with reactive strategies $S_1\!=\!(y_1, p_1, q_1$) and $S_2\!=\!(y_2,p_2,q_2)$ who interact in a repeated prisoner's dilemma with continuation probability $\delta$. We consider the vector 
\begin{equation}
\mathbf{v}(S_1,S_2)\!=\!\Big(v_{R}(S_1,S_2),v_{S}(S_1,S_2),v_{T}(S_1,S_2),v_{P}(S_1,S_2)\Big)\!:=\!(1\!-\!\delta)\mathbf{v_0}(I_4-\delta M)^{-1}.
\end{equation}
Here, $\mathbf{v_0}$ denotes the expected distribution of the four outcomes in the very first round, $I_4$ is the $4\!\times\!4$ identity matrix, and $M$ is the transition matrix of the Markov chain. 
The entries of $\mathbf{v}$ can be calculated explicitly,
\begin{equation} \label{Eq:LastRound}
\setlength{\arraycolsep}{1pt}
\begin{array}{rcl}

@@

%x_1(S_1,S_2)&=&\displaystyle\frac{(1\!-\!\delta)y_1+\delta\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big)}{1-\delta^2r_1r_2}\\[0.7cm]
%x_2(S_1,S_2)&=&\displaystyle\frac{(1\!-\!\delta)y_2+\delta\Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}{1-\delta^2r_1r_2}.\\
\end{array}
\end{equation}
In these expressions, we have used the notation $r_i:=p_i\!-\!q_i$, $\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$, and $\bar{r}_i:=\bar{p}_i\!-\!\bar{q}_i=-r_i$ for $i\!\in\!\{1,2\}$. Let $\mathcal{U}=\{R,S,T,P\}$ denote the set of feasible payoffs in each round, and let $\mathbf{u}\!=\!(R,S,T,P)$ be the corresponding payoff vector. Then one can show the following result.

\begin{Prop}
Consider a repeated prisoner's dilemma between players with reactive strategies $S_1\!=\!(y_1, p_1, q_1$)  and $S_2\!=\!(y_2,p_2,q_2)$ respectively. Then the probability that the $S_1$ player receives the payoff $u\!\in\! \mathcal{U}$ in the very last round of the game is given by $v_{u}(S_1,S_2)$, as given by Eq.~(\ref{Eq:LastRound}). 
\end{Prop}

@@

We compare this process for stochastic payoff evaluation with the analogous process where players update their strategies with respect to their {\it expected} payoffs,
\begin{equation} \label{Eq:ExpPay}
\begin{array}{lcrcr}
\displaystyle \pi_1	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(S_1,S_1),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(S_1,S_2),\mathbf{u}\rangle,\\[0.5cm]
\displaystyle \pi_2	&=	&\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(S_2,S_1),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(S_2,S_2),\mathbf{u}\rangle.\\
\end{array}
\end{equation}
In the limit of no discounting, $\delta\!\rightarrow\! 1$, this process based on expected payoffs has been considered in~\cite{imhof:prsb:2010}.

@@

Analogously, we can calculate the probability that the number of mutants decreases by one in one step of the evolutionary process. This probability is
\begin{equation}
\lambda^-_k=\frac{N\!-\!k}{N}\cdot\frac{k}{N} \sum_{u_1,u_2\in\mathcal{U}} x(u_1,u_2)\cdot \rho(u_2,u_1).
\end{equation}
The fixation probability of the mutant strategy then takes the standard form \cite{nowak:Nature:2004},
\begin{equation}
\varphi = \frac{1}{1+\sum_{i=1}^{N-1}\prod_k^i \frac{\lambda^-_k}{\lambda^+_k}}.
\end{equation}

@@

\subsection{Simulation results}

\FigEvoProc -- \FigResultsOverPara~show simulation results for the above described process. \FigEvoProc~depicts the evolving conditional cooperation probabilities $p$ and $q$ (assuming that the discount factor~$\delta$ and the benefit $b$ are comparably high). The left panel corresponds to the standard scenario considered in the previous literature. It considers players who use expected payoffs (\ref{Eq:ExpPay}) to update their strategies. The right panel shows the scenario considered herein, in which players update their strategies based on their last round's payoff. The figure suggests that when updating is based on expected payoffs, players tend to be more generous (their $q$-values are higher on average). In addition, players are generally more cooperative.

\begin{figure}[t!]
\centering
\includegraphics[width=0.75\textwidth]{Fig1} 
\caption{{\bf Evolutionary dynamics under expected payoffs (left) and stochastic payoffs (right).} 
We have run two simulations of the evolutionary process described in Section~\ref{Sec:BasicSetup} for $T\!=\!10^7$ time steps. For each time step, we have recorded the current resident population ($y,p,q$). Since simulations are run for a relatively high continuation probability of $\delta\!=\!0.999$, we do not report the players' initial cooperation probability $y$. The graphs show how often the resident population chooses each combination ($p,q$) of conditional cooperation probabilities in the subsequent rounds. ({\bf A}) If players update based on their expected payoffs, the resident population typically applies a strategy for which $p\!\approx\!1$ and $q\!\le\!1\!-\!c/b\!=\!0.9$. The cooperation rate within the resident population (averaged over all games and over all time steps) is close to 100\%. ({\bf B}) When players update their strategies based on their realized payoffs in the last round, there are two different predominant behaviors. The resident population either consists of defectors (with $p\!\approx\!q\!\approx\!0$) or of conditional cooperators. In the latter case, the maximum level of $q$ consistent with stable cooperation is somewhat smaller compared to the expected-payoff setting, $q\!<\!0.5$. Also the resulting cooperation rate is smaller. On average, players cooperate roughly in half of all rounds. Parameters: $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.999$.}
\end{figure}

To obtain some intuition for this result, we have recorded which mutant strategies typically invade into an ALLD population, and which mutants invade into a GTFT population, for each of the two scenarios (\FigInvAnalysis). Compared to expected-payoff updating, the stochastic scenario drastically reduces the stability of GTFT. On average, it takes fewer mutants until a GTFT population is successfully invaded. Moreover, successful mutants  tend to be less cooperative.

\begin{figure}[t!]
\centering
\includegraphics[width=0.75\textwidth]{Fig2} ~\\[-0.6cm]
\caption{{\bf Invasion into ALLD and into GTFT for the two evolutionary scenarios.} 
For this figure, we either consider a resident population of ALLD players (top) or of GTFT players (bottom). In each case, we have run 1,000 independent simulations of the evolutionary process described in Section~\ref{Sec:BasicSetup}. The colored dots within the four panels depict which mutant strategies successfully invaded into the respective resident population. The black arrow indicates the average over all successful mutant strategies. In addition, we have recorded how many mutant strategies need to be introduced into the population until the resident becomes replaced. Stochastic payoffs tend to increase the time until ALLD is invaded, and they reduce the time until GTFT is invaded. Parameters are the same as in \FigEvoProc. Strategies are defined as $ALLD\!=\!(0,0,0)$ and $GTFT=(1,1,1/3)$.}
~\\[0.4cm]
\includegraphics[width=0.8\textwidth]{Fig3} 
\caption{{\bf The evolution of cooperation for different parameter values.} 
While the previous figures depict the evolutionary outcome for fixed parameter values, here we vary the benefit of cooperation $b$, the strength of selection $\beta$, and the expected number of rounds, $1/(1\!-\!\delta)$. In all cases, stochastic payoff evaluation tends to reduce the evolving cooperation rates. Unless explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for $T\!=\!5\times 10^6$ time steps for each parameter combination.}
\end{figure}

Finally, we have also explored how the evolving cooperation rates change as we vary the benefit $b$, the selection strength $\beta$, and the discount factor $\delta$ (\FigResultsOverPara). In all three cases, we find that the two scenarios yield similar cooperation rates when the respective parameters are small. Once there is a high benefit to cooperation, strong selection, or a high expected number of rounds, updating based on expected payoffs yields higher cooperation rates.

@@

{\setlength{\bibsep}{0\baselineskip} \small
\bibliographystyle{unsrt}
\bibliography{\string~/Dropbox/Bibliography/ET}}

>>>>

\documentclass[11pt]{article}
\usepackage{times}
\usepackage{amsmath,amsthm,amssymb,setspace,enumitem,epsfig,titlesec,verbatim,color,array,eurosym,multirow}
\usepackage[sort&compress]{natbib}
\usepackage[footnotesize,bf]{caption}
\usepackage[margin=2.5cm, includefoot, footskip=30pt]{geometry}
\usepackage{standalone}
\usepackage{tikz}
\smallskip % Erlaubt kleine Abstaende zwischen Paragraphen, falls es dem Seitenlayout hilft
\renewcommand{\baselinestretch}{1.3}
\newcommand{\R}{\mathbb{R}}

@@

\section{Introduction}

Evolutionary game theory consists of two main approaches. The first approach
derives from the work of Maynard Smith and Price and employs the concept of an
evolutionarily stable strategy. The second approach, constructs an explicit
model of the process by which the frequency of strategies change in the
population and studies properties of the evolutionary dynamics within that
model.

The frequency of the strategies depends on a strategy's fitness, and commonly,
the fitness of a strategy in a population is the expected payoff it receives
from an interaction with a random member of the population.

The aim of this work is to understand the effect of the fitness function on the
evolution of the population. More specifically, this works evaluates a number of
different approaches for estimating the fitness of an individual.

\section{Model}

The Prisoner's Dilemma (PD) is a two player no-cooperative game. The players
simultaneously and independently make a decision to Cooperate (C) or to defect
(D). The payoffs for both players depend on their actions and the actions of
their opponents. More specifically, the payoffs are given by:

\begin{equation}\label{eq:pd_normal_form}
    \begin{pmatrix}
        R & S  \\
        T & P
    \end{pmatrix}
\end{equation}

The payoffs, \((R, P, S, T)\), are constrained by \(T > R > P > S\) and \(2R > T + S\).

A special case is the donation game,” where each player can cooperate by providing 
a benefit \(b\) to the other player at their cost \(c\), with \(0 < c < b\).
Then, \(T=b, R=b-c, S=-c, P=0\), and matrix~(\ref{eq:pd_normal_form}) is give by:

\begin{equation}\label{eq:donation_normal_form}
    \begin{pmatrix}
        b-c & -c  \\
        b & 0
    \end{pmatrix}
\end{equation}

where \(c\) is the cost inflicted to an individual for cooperating and \(b\) is
the benefit.

The questions of evolution become more interesting when repetition is considered,
and the history can be accessed when making decisions. The iterated form of the
game is called the Iterated Prisoner's Dilemma (IPD). There are several two
player repeated games such as the Snowdrift game~\cite{sugden2004economics},
Harmony and the Stag Hunt~\cite{skyrms2001stag} game. We present results in all
four games.

The space of strategies in the IPD is infinite. To continue with our
evolutionary analysis, we assume herein that individuals at most make use of
simple \textbf{reactive strategies}. Reactive strategy are a set of memory-one
strategies that only take into account the previous action of the opponent. An
example of a reactive strategy is Tit For Tat. Reactive strategies can be
written explicitly as a vector \(\in \R_{3}\). More specifically a reactive
strategy \(s\) is given by \(s=(y, p, q)\) where \(y\) is the probability that
the strategy opens with a cooperation and \(p, q\) are the probabilities that
the strategy cooperates given that the opponent cooperated and defected
equivalently.

A match between two reactive strategies takes the form of a stochastic process,
and more specifically, of a Markov chain with four possible states \(CC, CD, DC,
DD\) (the possible outcomes of each round). We assume a match between two
reactive strategies \(s_{1}=(y_{1}, p_{1}, q_{1})\) and \(s_{2}=(y_{2}, p_{2},
q_{2})\), the Markov process is described by the transition matrix \(M\):

\begin{equation}
    \input{static/matrix}
\end{equation}

The long run steady state probability vector \(\mathbf{v}\), which is the
solution to \(\mathbf{v} M = \mathbf{v}\), can be combined with the payoff of
matrix (\ref{eq:pd_normal_form}) (denoted as \(U\)) to give the expected payoffs
for each player. More specifically, the payoffs for a reactive strategy \(s_1\)
against an opponent \(s_2\) is:

\begin{equation}
    \mathbf{v}(s_1, s_2) \cdot U
\end{equation}

where

\begin{equation}
    U = \{R, S, T, P\}.
\end{equation}

\subsection{Evolutionary Dynamics}

In evolution context we consider a population of \(N\) players,
where \(N\) is even, and where mutations are sufficiently rare. At any point
in time the there are at most two different strategies are present in
the population. Suppose there are \(N - k\) players who use the strategy \(s_{1}=(y_{1},
p_{1}, q_{1})\), whereas \(k\) players use the strategy \(s_{2}=(y_{2}, p_{2},
q_{2})\). We refer to these two player types as ‘residents’ and ‘mutants’,
respectively.

Each step of the evolutionary process consists of two stages, a game stage and
an updating stage.

\begin{enumerate}
    \item In the game stage, each player is randomly matched with some
    other player in the population to interact in one instance of the IPD.
    \item In the updating stage, two players are randomly drawn from
    the population, a `learner' and a `exemplar'. Given that the learner's payoff
    in the last round is $u_L\!\in\! \mathcal{U}$ and that the exemplar's last
    round's payoff $u_E\!\in\! \mathcal{U}$, we assume the learner adopts the role
    model's strategy with probability

    \begin{equation} \label{Eq:rho}
    \rho(u_L, u_E) = \frac{1}{1\!+\!\exp\big[ \!-\!\beta (u_E\!-\!u_L) \big]}. 
    \end{equation}

    where $\beta\!\ge\!0$ corresponds to the strength of selection.
\end{enumerate}

We iterate this basic evolutionary step until either the mutant strategy goes
extinct, or until it fixes in the population (in which case the mutant strategy
becomes the new resident strategy). After either outcome, we introduce a new
mutant strategy $s'_2\!=\!(y_2',p_2',q_2')$ (uniformly chosen from all reactive
strategies at random), and we set the number of mutants to $k\!=\!1$. This
process of mutation and fixation/extinction is then iterated many times.

We compare this process for what we defined as {\bf stochastic payoff
evaluation} with the analogous process where players update their strategies
with respect to their {\bf expected} payoffs,

\begin{equation} \label{eq:ExpPay}
\begin{array}{lcrcr}
\pi_1 &= &\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_1,s_1),\mathbf{U}\rangle	& + &\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_1,s_2),\mathbf{U}\rangle,\\[0.5cm]
\pi_2 &= &\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_2,s_1),\mathbf{U}\rangle & + &\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_2,s_2),\mathbf{U}\rangle.\\
\end{array}
\end{equation}

In the limit of no discounting, $\delta\!\rightarrow\! 1$, this process based on
expected payoffs has been considered in~\cite{imhof2010stochastic}.

\subsection{Stochastic payoff evaluation}

We define {\bf stochastic payoff} as the average payoff $u\!\in\! \mathcal{U}$ a player
receives in the last \(n\) rounds of the game given that they interact with
\(m\) players.

\subsubsection*{Case \(n=m=1\).}

Initially, consider the situation where \(n=m=1\). The player's stochastic
payoff is what they receive in the last round against a single opponent. There
only four possible outcomes for the last round, those are \(CC, CD, DC, DD\).
Consider two players with reactive strategies $S_1\!=\!(y_1, p_1, q_1$) and
$S_2\!=\!(y_2,p_2,q_2)$ who interact in a repeated prisoner's dilemma with
continuation probability $\delta$, the probability that are in each of the
four possible states in the last round is given by:

\begin{equation}
    \mathbf{v}(s_1,s_2)\!=\!\Big(\mathbf{v}_{R}(s_1,s_2),\mathbf{v}_{S}(s_1,s_2),\mathbf{v}_{T}(s_1,s_2),\mathbf{v}_{P}(s_1,s_2)\Big).
\end{equation}

where,

\begin{equation} \label{Eq:LastRound}
    \setlength{\arraycolsep}{1pt}
    \begin{array}{rcl}

    \mathbf{v}_{R}(S_1,S_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
    {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]

    \mathbf{v}_{S}(S_1,S_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
    {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]

    \mathbf{v}_{T}(S_1,S_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
    {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]

    \mathbf{v}_{P}(S_1,S_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
    {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)}.
    \end{array}
\end{equation}

\begin{proof}

Assume a repeated prisoner's dilemma between two reactive strategies. Given the
continuation probability $\delta$, probability that the game ends in the after
the first round $(1 - \delta)$ and the expected distribution of the four
outcomes in the very first round is $\mathbf{v_0}$ defined as. Following the
first round the, the outcome of the next rounds with a probability $\delta$
is \(M\) such that,

\[\dots\]

It can shown that, \(!(1\!-\!\delta)\mathbf{v_0}(I_4-\delta M)^{-1}\) and with
some algebraic manipulation we derive to Equation~\ref{Eq:LastRound}.

\end{proof}

% \begin{Prop}
%     Consider a repeated prisoner's dilemma, with
%     continuation probability $\delta$, between players with reactive strategies
%     $s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Then the
%     probability that the $s_1$ player receives the payoff $u\!\in\! \mathcal{U}$ in
%     the very last round of the game is given by $v_{u}(S_1,S_2)$, as given by
%     Eq.~(\ref{Eq:LastRound}).
% \end{Prop}

is given by $v_{u}(S_1,S_2)$

Consider two players with reactive strategies $S_1\!=\!(y_1, p_1, q_1$) and $S_2\!=\!(y_2,p_2,q_2)$ who interact in a repeated prisoner's dilemma with continuation probability $\delta$. We consider the vector

\begin{equation}
\mathbf{v}(S_1,S_2)\!=\!\Big(v_{R}(S_1,S_2),v_{S}(S_1,S_2),v_{T}(S_1,S_2),v_{P}(S_1,S_2)\Big)\!:=\!(1\!-\!\delta)\mathbf{v_0}(I_4-\delta M)^{-1}.
\end{equation}

Here, $\mathbf{v_0}$ denotes the expected distribution of the four outcomes in the very first round, $I_4$ is the $4\!\times\!4$ identity matrix, and $M$ is the transition matrix of the Markov chain. 
The entries of $\mathbf{v}$ can be calculated explicitly,

\begin{equation} \label{Eq:LastRound}
\setlength{\arraycolsep}{1pt}
\begin{array}{rcl}

@@

%x_1(S_1,S_2)&=&\displaystyle\frac{(1\!-\!\delta)y_1+\delta\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big)}{1-\delta^2r_1r_2}\\[0.7cm]
%x_2(S_1,S_2)&=&\displaystyle\frac{(1\!-\!\delta)y_2+\delta\Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}{1-\delta^2r_1r_2}.\\
\end{array}
\end{equation}

In these expressions, we have used the notation $r_i:=p_i\!-\!q_i$, $\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$, and $\bar{r}_i:=\bar{p}_i\!-\!\bar{q}_i=-r_i$ for $i\!\in\!\{1,2\}$. Let $\mathcal{U}=\{R,S,T,P\}$ denote the set of feasible payoffs in each round, and let $\mathbf{u}\!=\!(R,S,T,P)$ be the corresponding payoff vector. Then one can show the following result.

\includestandalone[width=0.35\textwidth]{static/expected_stochastic}

\includestandalone{static/stochastic}

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=\textwidth]{static/expected_and_stochastic_for_donation_game.pdf}
    \caption{Blah blah.}
\end{figure}

@@

We compare this process for stochastic payoff evaluation with the analogous process where players update their strategies with respect to their {\it expected} payoffs,
\begin{equation} \label{Eq:ExpPay}
\begin{array}{lcrcr}
\displaystyle \pi_1	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(S_1,S_1),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(S_1,S_2),\mathbf{u}\rangle,\\[0.5cm]
\displaystyle \pi_2	&=	&\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(S_2,S_1),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(S_2,S_2),\mathbf{u}\rangle.\\
\end{array}
\end{equation}
In the limit of no discounting, $\delta\!\rightarrow\! 1$, this process based on expected payoffs has been considered in~\cite{imhof2010stochastic}.

@@

Analogously, we can calculate the probability that the number of mutants decreases by one in one step of the evolutionary process. This probability is
\begin{equation}
\lambda^-_k=\frac{N\!-\!k}{N}\cdot\frac{k}{N} \sum_{u_1,u_2\in\mathcal{U}} x(u_1,u_2)\cdot \rho(u_2,u_1).
\end{equation}
The fixation probability of the mutant strategy then takes the standard form \cite{nowak2004emergence},
\begin{equation}
\varphi = \frac{1}{1+\sum_{i=1}^{N-1}\prod_k^i \frac{\lambda^-_k}{\lambda^+_k}}.
\end{equation}

@@

% \subsection{Simulation results}

% \FigEvoProc -- \FigResultsOverPara~show simulation results for the above described process. \FigEvoProc~depicts the evolving conditional cooperation probabilities $p$ and $q$ (assuming that the discount factor~$\delta$ and the benefit $b$ are comparably high). The left panel corresponds to the standard scenario considered in the previous literature. It considers players who use expected payoffs (\ref{Eq:ExpPay}) to update their strategies. The right panel shows the scenario considered herein, in which players update their strategies based on their last round's payoff. The figure suggests that when updating is based on expected payoffs, players tend to be more generous (their $q$-values are higher on average). In addition, players are generally more cooperative.

% \begin{figure}[t!]
% \centering
% \includegraphics[width=0.75\textwidth]{Fig1} 
% \caption{{\bf Evolutionary dynamics under expected payoffs (left) and stochastic payoffs (right).} 
% We have run two simulations of the evolutionary process described in Section~\ref{Sec:BasicSetup} for $T\!=\!10^7$ time steps. For each time step, we have recorded the current resident population ($y,p,q$). Since simulations are run for a relatively high continuation probability of $\delta\!=\!0.999$, we do not report the players' initial cooperation probability $y$. The graphs show how often the resident population chooses each combination ($p,q$) of conditional cooperation probabilities in the subsequent rounds. ({\bf A}) If players update based on their expected payoffs, the resident population typically applies a strategy for which $p\!\approx\!1$ and $q\!\le\!1\!-\!c/b\!=\!0.9$. The cooperation rate within the resident population (averaged over all games and over all time steps) is close to 100\%. ({\bf B}) When players update their strategies based on their realized payoffs in the last round, there are two different predominant behaviors. The resident population either consists of defectors (with $p\!\approx\!q\!\approx\!0$) or of conditional cooperators. In the latter case, the maximum level of $q$ consistent with stable cooperation is somewhat smaller compared to the expected-payoff setting, $q\!<\!0.5$. Also the resulting cooperation rate is smaller. On average, players cooperate roughly in half of all rounds. Parameters: $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.999$.}
% \end{figure}

% To obtain some intuition for this result, we have recorded which mutant strategies typically invade into an ALLD population, and which mutants invade into a GTFT population, for each of the two scenarios (\FigInvAnalysis). Compared to expected-payoff updating, the stochastic scenario drastically reduces the stability of GTFT. On average, it takes fewer mutants until a GTFT population is successfully invaded. Moreover, successful mutants  tend to be less cooperative.

% \begin{figure}[t!]
% \centering
% \includegraphics[width=0.75\textwidth]{Fig2} ~\\[-0.6cm]
% \caption{{\bf Invasion into ALLD and into GTFT for the two evolutionary scenarios.} 
% For this figure, we either consider a resident population of ALLD players (top) or of GTFT players (bottom). In each case, we have run 1,000 independent simulations of the evolutionary process described in Section~\ref{Sec:BasicSetup}. The colored dots within the four panels depict which mutant strategies successfully invaded into the respective resident population. The black arrow indicates the average over all successful mutant strategies. In addition, we have recorded how many mutant strategies need to be introduced into the population until the resident becomes replaced. Stochastic payoffs tend to increase the time until ALLD is invaded, and they reduce the time until GTFT is invaded. Parameters are the same as in \FigEvoProc. Strategies are defined as $ALLD\!=\!(0,0,0)$ and $GTFT=(1,1,1/3)$.}
% ~\\[0.4cm]
% \includegraphics[width=0.8\textwidth]{Fig3} 
% \caption{{\bf The evolution of cooperation for different parameter values.} 
% While the previous figures depict the evolutionary outcome for fixed parameter values, here we vary the benefit of cooperation $b$, the strength of selection $\beta$, and the expected number of rounds, $1/(1\!-\!\delta)$. In all cases, stochastic payoff evaluation tends to reduce the evolving cooperation rates. Unless explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for $T\!=\!5\times 10^6$ time steps for each parameter combination.}
% \end{figure}

% Finally, we have also explored how the evolving cooperation rates change as we vary the benefit $b$, the selection strength $\beta$, and the discount factor $\delta$ (\FigResultsOverPara). In all three cases, we find that the two scenarios yield similar cooperation rates when the respective parameters are small. Once there is a high benefit to cooperation, strong selection, or a high expected number of rounds, updating based on expected payoffs yields higher cooperation rates.

@@

\bibliographystyle{unsrt}
\bibliography{bibliography.bib}
